# Diagnostic cardinality

## Overview

```{r}
# Relative location of data.
working_directory <-
  "experiments/2021-05-27-cardinality/analysis/"
# working_directory <- "./"

# Settings for visualization
cb_palette <- "Set2"
# Create directory to dump plots
dir.create(paste0(working_directory, "imgs"), showWarnings=FALSE)
```

## Analysis dependencies

```{r, message=FALSE}
library(ggplot2)
library(tidyverse)
library(cowplot)
library(viridis)
library(RColorBrewer)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

These analyses were conducted in the following computing environment:

```{r}
print(version)
```


## Setup

```{r}
data_loc <- paste0(
  working_directory,
  "data/timeseries-res-1000g.csv"
)
data <- read.csv(
  data_loc,
  na.strings="NONE"
)

data$cardinality <- as.factor(
  data$OBJECTIVE_CNT
)
data$selection_name <- as.factor(
  data$selection_name
)

data$elite_trait_avg <-
  data$ele_agg_per / data$OBJECTIVE_CNT

data$unique_start_positions_coverage <-
  data$uni_str_pos / data$OBJECTIVE_CNT

####### misc #######
# Configure our default graphing theme
theme_set(theme_cowplot())
```

## Exploration diagnostic performance

First, we look at performance over time.
Specifically, we look at the normalized aggregage score of the most performant individuals over time.
To control for different cardinalities having different maximum scores, we normalized performances (by dividing by cardinality) to values between 0 and 100.

```{r}
elite_trait_ave_fit <- ggplot(
    data,
    aes(
      x=gen,
      y=elite_trait_avg,
      color=cardinality,
      fill=cardinality
    )
  ) +
  stat_summary(geom="line", fun=mean) +
  stat_summary(
    geom="ribbon",
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    alpha=0.2,
    linetype=0
  ) +
  scale_y_continuous(
    name="Average trait performance",
    limits=c(0, 100)
  ) +
  scale_x_continuous(
    name="Generation"
  ) +
  scale_fill_brewer(
    name="Cardinaltiy",
    palette=cb_palette
  ) +
  scale_color_brewer(
    name="Cardinaltiy",
    palette=cb_palette
  ) +
  ggsave(
    paste(working_directory, "imgs/elite_trait_average_ot.pdf", sep="")
  )
elite_trait_ave_fit
```

### Final performance

Next, we look only at the final performances of each treatment

```{r}
final_data <- filter(data, gen==max(data$gen))
elite_trait_ave_fit_final <- ggplot(
    final_data,
    aes(x=cardinality, y=elite_trait_avg, fill=cardinality)
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8,
    scale="width"
  ) +
  geom_point(
    mapping=aes(color=cardinality),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Average trait performance",
    limits=c(0, 100)
  ) +
  scale_x_discrete(
    name="Cardinality"
  ) +
  scale_fill_brewer(
    name="Cardinaltiy",
    palette=cb_palette
  ) +
  scale_color_brewer(
    name="Cardinaltiy",
    palette=cb_palette
  ) +
  theme(
    legend.position="none"
  )
elite_trait_ave_fit_final
```

## Unique starting positions

Next, we analyze the number of unique starting position maintained by populations.

```{r}
ggplot(data, aes(x=gen, y=uni_str_pos, color=cardinality)) +
  stat_summary(geom="line", fun=mean) +
  stat_summary(
    geom="ribbon",
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    alpha=0.2,
    linetype=0
  ) +
  scale_y_continuous(
    name="Unique starting positions (population)",
  ) +
  scale_x_continuous(
    name="Generation"
  ) +
  ggsave(
    paste(
      working_directory,
      "imgs/pop_unique_starting_positions_ot.pdf",
       sep=""
    )
  )
```

Different cardinalities have numbers of possible starting positions, so next, we look at the proportion of starting positions (out of all possible) maintained by populations.

```{r}
unique_start_positions_coverage_fig <- ggplot(
    data,
    aes(
      x=gen,
      y=unique_start_positions_coverage,
      color=cardinality,
      fill=cardinality
    )
  ) +
  stat_summary(geom="line", fun=mean) +
  stat_summary(
    geom="ribbon",
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    alpha=0.2,
    linetype=0
  ) +
  scale_y_continuous(
    name="Starting position coverage",
    limits=c(0.0, 1.05)
  ) +
  scale_x_continuous(
    name="Generation"
  ) +
  scale_fill_brewer(
    name="Cardinaltiy",
    palette=cb_palette
  ) +
  scale_color_brewer(
    name="Cardinaltiy",
    palette=cb_palette
  ) +
  ggsave(
    paste(
      working_directory,
      "imgs/pop_unique_starting_position_coverage_ot.pdf",
      sep=""
    )
  )
unique_start_positions_coverage_fig
```

### Final starting position coverage

```{r}
final_unique_start_positions_coverage_fig <- ggplot(
    final_data,
    aes(
      x=cardinality,
      y=unique_start_positions_coverage,
      fill=cardinality
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8,
    scale="width"
  ) +
  geom_point(
    mapping=aes(color=cardinality),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Starting position coverage",
    limits=c(0, 1.05)
  ) +
  scale_x_discrete(
    name="Cardinality"
  ) +
  scale_fill_brewer(
    name="Cardinaltiy",
    palette=cb_palette
  ) +
  scale_color_brewer(
    name="Cardinaltiy",
    palette=cb_palette
  ) +
  theme(
    legend.position="none"
  )
final_unique_start_positions_coverage_fig
```

## Manuscript figures

Combine figures for the manuscript.

```{r}
grid <- plot_grid(
  elite_trait_ave_fit +
    ggtitle("Performance over time") +
    theme(legend.position="none"),
  elite_trait_ave_fit_final +
    ggtitle("Final performance") +
    theme(),
  unique_start_positions_coverage_fig +
    ggtitle("Start position coverage over time") +
    guides(color=guide_legend(nrow = 1), fill=guide_legend(nrow=1)) +
    theme(
      legend.position="bottom",
      legend.box="horizontal"
    ),
  final_unique_start_positions_coverage_fig +
    ggtitle("Final start position coverage") +
    theme(),
  nrow=2,
  ncol=2,
  rel_widths=c(2,1),
  labels="auto"
)

save_plot(
  paste(working_directory, "imgs/cardinality-panel.pdf", sep=""),
  grid,
  base_width=12,
  base_height=10
)

grid
```